On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-19 paper.tex
Writing section on Rehmann theorem

\section{A variant of Rehmann theorem}
Let $A$ be arbitrary commutative unital ring and $\mathfrak{m}$ be its ideal.
Denote by $B$ the ring $\mathfrak{m}[t^{-1}] + A[t]$ with the obvious $\mathbb{Z}$-grading.
Clearly, the $k$-th homogeneous component $B_k$ of $B$ equals $A\cdot t^k$ for $k \geq 0$ or $B_k = \mathfrak{m} \cdot t^k$ for $k<0$.
If $a$ is a homogeneous element of $B$ we denote its homogeneous degree by $|a|$.

In what follows we will be concerned with finding a presentation of the Steinberg group $\St(4, B)$ in terms of those generators $x_{ij}(\xi)$
 for which $\xi$ is homogeneous of degree $\leq 1$ (i.\,e. $|\xi| \leq 1$).
Our presentation should be thought of as an extension of~\cite[Lemma~3.3]{Tu83} to the case $n=4$ (with variables $t$ and $t^{-1}$ swapped).
Alternatively, our result can be thought of as a variant of the classical theorem of Rehmann~\cite{Re75} (cf. also with Rehmann--Soul{\'e} theorem in~\cite{RS76}).

There are several reasons why we can not simply refer to~\cite{Re75} or~\cite{RS76} and need to repeat arguments from these papers.
\begin{itemize}
 \item Rehmann and Soul{\'e} deal with usual polynomial rings rather than $\mathbb{Z}$-graded rings
  (i.\,e. their result corresponds to the special case $\mathfrak{m}=0$ of our result);
 \item their presentation is formulated in terms of generators $x_{ij}(\xi)$ with $|\xi| \leq 2$ while for our purposes we need a presentation in terms
  of the smaller set of generators with $| \xi |\leq 1$;
 \item it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}, while work with
  arbitrary coefficient ring $A$.
\end{itemize}

For $m \geq 1$ consider the following set of generators:
\[\XX{m} = \{ x_{ij}(\xi) \mid \xi \in B_k,\text{ for } k\leq m,\ 1\leq i, j\leq 4,\ i\neq j\}. \]
It is obvious that $\XX{m} \subseteq \XX{m+1}$, denote by $\XX{\infty}$ the union of all $\XX{n}$'s.

For $m\geq 1$ denote by $G_m$ the group presented by generators $\mathcal{X}_m$ and the set of relations $\RR{m}$ consisting of the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{ij}(\xi) x_{ij}(\eta) & = x_{ij}(\xi+\eta),&  \xi,\eta\in B_k,\ k\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\alpha'}(\eta))] &  = 1, & \text{in the case $\alpha+\alpha'\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_k,\ \eta \in B_l,\ l,k\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\alpha'}(\eta)] & = x_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi\eta), & \text{in the case $\alpha+\alpha'\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_k,\ \eta \in B_l,\ l, k, l+k \leq m.
 \end{align}

 %The second is that they prove their results only in the special case $B = A[t]$, moreover, they make assumptions about the ground ring $A$
%().
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

 \begin{lemma} In the group $G_1$ the following two families of relations are equivalent:
\begin{align} 
\nonumber \tag{$d_1$}
[[x_{\alpha}(a t), x_{\beta}(b t)],  x_{\gamma}(c)] & = [x_{\alpha + \beta}(a b t),  x_{\gamma}(c t)], & a,b,c \in A,\ \alpha + \beta + \gamma \in \Phi\\
\nonumber \tag{$df_1$}
[[x_{ij}(a t), x_{jk}(b t)],  x_{kl}(c)] & = [x_{ik}(a b t),  x_{kl}(c t)]\\
\nonumber \tag{$e_1$}
[[x_{\alpha}(at), x_{\beta}(bt)], x_{\gamma}(ct) ]& = 1, & \alpha + \beta \in \Phi,\ \alpha+\beta \perp \gamma \\
\nonumber \tag{$ef_1$}
[[x_{ij}(at), x_{jk}(bt)], x_{jl}(ct) ]& = 1
\end{align}
 \end{lemma}

It is clear that the natural embedding of generators $\mathcal{X}_n \subseteq \mathcal{X}_{n+1}$ induces
 a map $\theta_n\colon G_n \to G_{n+1}$.
The aim of this subsection is to demonstrate an analogue of~\cite[Lemma~3.3]{Tu83}. 
\begin{lemma}\label{lem:tul3.3}
 Let $\Phi$ be a simply-laced root systems of rank $\geq 3$. 
 Then $ $
 Moreover, the map $\theta_1\colon G_1/H \to G_2$ induced by  is isomorphic to $G_1$.
\end{lemma}
Unfortunately, the lemma cannot be proved by a slight modification of the original short argument of Tulenbayev.
 %for even in the linear case it works only for $\mathrm{rk}(\Phi)\geq 4$, and, in particular, is unsuitable in the case $\Phi=\mathsf{D}_\ell$.
Instead, we have to employ a varition of a much longer argument of U.~Rehmann and C.~Soul{\'e}, cf~\cite{Re75,RS76}.

>>>>

\section{A variant of Rehmann theorem}
Let $A$ be arbitrary commutative unital ring and $\mathfrak{m}$ be its ideal.
Denote by $B$ the ring $\mathfrak{m}[t^{-1}] + A[t]$ with the obvious $\mathbb{Z}$-grading.
In other words, as a module $B$ decomposes as a direct sum of its homogeneous components $\oplus_k B_k$,
 where each $B_k$ is either $A\cdot t^k$ for $k \geq 0$ or $\mathfrak{m} \cdot t^k$ for $k<0$.
%If $a$ is a homogeneous element of $B$ (i.\,e. an element of some $B_k$) we denote its homogeneous degree by $|a|$.

Let $\Phi$ be a reduced irreducible root system of rank $\geq 3$.
The aim of this section is to find a presentation of the Steinberg group $\St(\Phi, B)$ in terms of generators $x_{\alpha}(\xi)$
 for which $\xi$ is homogeneous of degree $\leq 1$, see~\cref{prop:tul3.3} below.
Our presentation should be thought of as an analogue of~\cite[Lemma~3.3]{Tu83} in the case $n=4$ (with variables $t$ and $t^{-1}$ swapped).
Notice that Tulenbaev's short proof of \cite[Lemma~3.3]{Tu83} which works for $\Phi=\rA_\ell$, $\ell\geq 4$ does not generalize to the case $\ell=3$, 
 so we are forced to recourse to a much longer argument inspired by the proof of Rehmann--Soul{\'e} theorem on
 finite presentation of Steinberg groups over polynomial rings (cf.~\cite{Re75, RS76}).

For $m \geq 1$ consider the following set of Steinberg generators:
\[\XX{m} = \{ x_{\alpha}(\xi) \mid \xi \in B_d,\text{ for } d\leq m,\ \alpha\in\Phi\}. \]
Clearly, $\XX{m} \subseteq \XX{m+1}$, denote by $\XX{\infty}$ the union of all $\XX{n}$'s.

\begin{rem}
There are several reasons why we can not simply refer to~\cite{Re75} or~\cite{RS76} and need to reproduce arguments from these papers.
\begin{itemize}
 \item Rehmann and Soul{\'e} deal with usual polynomial rings rather than $\mathbb{Z}$-graded rings
  (i.\,e. their results correspond to the special case $\mathfrak{m}=0$ of our result);
 \item Presentations from~\cite{Re75,RS76} are formulated in terms of generating set $\XX{2}$ which is not sufficient for our purposes
  (we need a presentation in terms of $\XX{1}$);
 \item it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}, while we work with
  arbitrary coefficient ring $A$.
\end{itemize} 
\end{rem}

For $m\geq 2$ define $G_m$ as the group presented by generators $\mathcal{X}_m$ and the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{\alpha}(\xi) x_{\alpha}(\eta) & = x_{\alpha}(\xi+\eta),&  \xi,\eta\in B_d,\ d\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\beta}(\eta))] &  = 1, & \text{in the case $\alpha+\beta\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_d,\ \eta \in B_e,\ d,e\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\beta}(\eta)] & = x_{\alpha+\beta}(N_{\alpha,\beta}\xi\eta), & \text{in the case $\alpha+\beta\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_d,\ \eta \in B_e,\ d, e, d+e \leq m;
 \end{align}
For $m=1$ define $G_1$ as the group presented by generators $\mathcal{X}_m$, three families of relations $a_1, b_1, c_1$ as above and the following extra family of relations:
\begin{align}  \label{eq:d1} \tag{$d_1$} [[x_{\alpha}(\xi), x_{\beta}(\eta)], x_{\gamma}(\zeta) ] & = 1, & \xi,\eta,\zeta\in B_1,\ \alpha + \beta \in \Phi,\ \alpha + \beta + \gamma \not\in \Phi\cup\{0\}.
\end{align}
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

It is clear that the natural embedding of generators $\mathcal{X}_n \subseteq \mathcal{X}_{n+1}$ induces
 a map $\theta_n\colon G_n \to G_{n+1}$.
Our main goal is to show the following result.
\begin{prop}\label{prop:tul3.3}
 For $n\geq 1$ the map $\theta_n$ is an isomorphism. Consequently, $\St(\Phi, B)$ can be presented by generators $\mathcal{X}_1$
  and four families of relations $a_1$, $b_1$, $c_1$, $d_1$.
\end{prop}

 \begin{lemma} In the group $G_1$ the following relation holds:
\begin{align} 
 [[x_{\alpha}(a t), x_{\beta}(b t)],  x_{\gamma}(c)] & = [x_{\alpha+\beta}(a b t),  x_{\gamma}(c t)].
\end{align}
 \end{lemma}
